---
title: "ghgtools-template"
author: "Brandon McNamara"
output: html_document
---
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(ggplot2)
library(kableExtra)
library(scales)
```

**ghgtools** is a project of the Carbon Accounting, Reporting, and Management lab [(CARML)](https://carml.rc.nau.edu/). The transition from a fossil fuel economy to a renewable energy economy is one of the most complex problems humanity has ever faced. Tracking, measuring, and reporting GHG emissions is essential to understand risks and identify opportunities related to climate change and the transition towards a renewable energy economy. Our mission is to grow the practice of GHG accounting with tools rooted in uncompromising transparency, rigorous data quality, and purposeful versatility. 

## Instructions

Navigate to the directory and open the following files. Enter your data. 

**AssetPortfolio** - The list of all your assets. This likely includes buildings and vehicles. You may also designate an _Enterprise_ asset to calculate scope 3 emissions for purchasing, business travel, and other supply chain related activities. 

**ActivityData** - The record of energy consumption and other GHG-producing activities across each of your assets. 

## How it Works

Most of the functionality in ghgtools comes from the [data.table](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html) package. 

```{r, include=TRUE, message=FALSE, warning=FALSE}
library(data.table)
```

Select your desired global warming potentials _(GWPS)_. You can learn more about global warming potentials by follwoing [this link](https://www.epa.gov/ghgemissions/understanding-global-warming-potentials). ghgtools defaults to the UNFCCC guidelines, which require the use of GWP values from the IPCC's Fifth Assessment Report (AR5).

```{r, include=TRUE, warning=FALSE}
GWP <- "AR5"
```

The emission factor library, abbreviated as EFL, is loaded from the ghgtools directory. The raw form of the EFL has a row of data for each greenhouse gas produced by the activity. For example, burning natural gas will generate CO2, CH4 and N2O. The following code loads the raw EFL and consolidates each activity into the CO2e emission factor using the selected GWPs. 

```{r, include=TRUE, warning=FALSE}
#load emission factor library

EFL <- fread("EFL.csv")
GWPs <- fread("GWPs.csv")
GWPs <- GWPs[, .(ghg,get(GWP))]
colnames(GWPs)[2] <- "GWP"
EFL_CO2e <- data.table(merge.data.table(EFL, GWPs, sort = FALSE, all.x = TRUE))
EFL_CO2e[, sum_co2e := ghg_emission_factor * GWP]
EFL_CO2e <- EFL_CO2e[, .(kgco2e_perunit = sum(sum_co2e)), by = .(ef_source, ef_publishdate, ef_activeyear, service_type, unit, emission_category, service_subcategory1, service_subcategory2, emission_scope, country, subregion)]
EFL_CO2e[, ef_publishdate := format(ef_publishdate, "%m/%d/%Y")]
EFL_CO2e[EFL_CO2e == ""] <- NA
setnames(EFL_CO2e, "ef_activeyear", "year")
```

<br>
Load your asset portfolio and activity data. **Make sure to replace the template data with your data**

```{r, echo=FALSE, warning=FALSE, message=FALSE}
AssetPortfolio <- fread("AssetPortfolio.csv")
ActivityData <- fread("ActivityData.csv")
```

<br>
Run the following code chunk to calculate GHG emissions for each record of activity data.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
eGRIDlookup <- fread("eGRID_lookup.csv")
Ecat_lookup <- fread("Ecat_lookup.csv")
DT1 <- fread("ActivityData.csv")
DT1[DT1 == ""] <- NA
DT2 <- data.table(merge.data.table(DT1, AssetPortfolio, sort = FALSE, all.x = TRUE))
DT3 <- data.table(merge.data.table(DT2, Ecat_lookup, by = c("asset_type", "service_type"), sort = FALSE))
DT4 <- data.table(merge.data.table(DT3, eGRIDlookup, sort = FALSE, all.x = TRUE))
GHGrawdata <- data.table(merge.data.table(DT4, EFL_CO2e, by = c("year", "service_type", "emission_category", "service_subcategory1", "service_subcategory2", "country", "subregion", "unit"), all.x = TRUE, sort = FALSE))
GHGrawdata[, kg_co2e := usage * kgco2e_perunit]
GHGrawdata[, MT_co2e := kg_co2e/1000]
setcolorder(GHGrawdata, c("asset_id", "asset_type", "asset_subtype", "address", "city", "state", "zip", "country", "region", "subregion", "business_unit", "year_built", "sqft", "service_type", "unit", "vendor", "account_id", "meter_number", "date", "year", "cost", "usage", "emission_category", "service_subcategory1", "service_subcategory2", "emission_scope", "kgco2e_perunit", "kg_co2e", "MT_co2e", "ef_source", "ef_publishdate"))
fwrite(GHGrawdata, "GHGrawdata.csv")
```

<br>
Here are a few example charts to get started with data visualization.

```{r, echo=FALSE, message = FALSE, warning=FALSE}
ghg_summary <- GHGrawdata[, .(EmissionTotal = sum(MT_co2e)), by = emission_scope]
GHG_sum_chart <- ggplot(ghg_summary, aes(x = emission_scope, y = EmissionTotal)) +
  geom_bar(stat = "identity", fill = "darkgreen", color = "black") +
  labs(title = "GHG Emissions Total",
       x = "",
       y = "MT CO2e")
print(GHG_sum_chart)
```

<br>

```{r, echo=FALSE, message = FALSE, warning=FALSE}
Ecat_summary <- GHGrawdata[, .(EmissionTotal = sum(MT_co2e)), by = emission_category]
Ecat_sum_chart <- ggplot(Ecat_summary, aes(x = reorder(emission_category, EmissionTotal), y = EmissionTotal)) +
  geom_bar(stat = "identity", fill = "darkgreen", color = "black") +
  labs(title = "GHG Emissions by Category",
       x = "",
       y = "MT CO2e") +
  coord_flip()
print(Ecat_sum_chart)
```

<br>

```{r, echo=FALSE, message = FALSE, warning=FALSE}
scope1and2 <- c("Scope 1", "Scope 2")
state_data <- GHGrawdata[emission_scope %in% scope1and2, .(EmissionTotal = sum(MT_co2e)), by = state]
state_data_chart <- ggplot(state_data, aes(x = reorder(state, EmissionTotal), y = EmissionTotal)) +
  geom_bar(stat = "identity", fill = "darkgreen", color = "black") +
  labs(title = "GHG Emissions by State",
       x = "",
       y = "MT CO2e") +
  coord_flip()
print(state_data_chart)
```
